DDSAnalytics  - Employee Trends

### 2018-04-16 msds6371-group-project-kaggle.Rmd
latest code output and images

---
title: "DDSAnalytics  - Employee Trends"
author:
- Randall Hendrickson
- Rajat Chandna
- Lokesh Maganti
- Mike Kelleher
date: "April 15, 2018"
output: 
  pdf_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
  html_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
---

@@

>>>>

---
title: "DDSAnalytics  - Employee Trends"
author:
- Randall Hendrickson
- Rajat Chandna
- Lokesh Maganti
- Mike Kelleher
date: "April 15, 2018"
output: 
  html_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
  pdf_document:
    toc: true
    toc_depth: 2
    fig_caption: yes
---

@@

### Training Code

```

/* 
Data Sets: 
- Train0 - Kaggle Training Data Set 
- TrainRed - Kaggle Training Data Set Reduced for question 1  
*/

filename CSV URL "https://dl.dropboxusercontent.com/spa/afq05cp80hp4ezn/downloads/public/msds-data/train.csv";

/* **************************************** */
/* DATA CLEANUP - use proc means to show missing values*/
/* **************************************** */

title "Train0 - Data Set Missing Values";
proc means data=train0 nmiss n; run;

title "Train0 - First 10";
proc print data=train0 (obs=10);
run;
title "";

/* **************************************** */
/* END DATA CLEANUP - after second proc means, no missing values */
/* **************************************** */

/******************************/
/** train0 is for Question 2 **/
/******************************/
proc sgscatter data=train0;
  plot SalePrice*GrLivArea /group=Neighborhood;
run;

/******************************/
/** train0 is for Question 1 **/
/******************************/
data trainRed;
  set train0 (keep=Id Neighborhood GrLivArea SalePrice);

  select (Neighborhood);
    when ("NAmes") Nei=Neighborhood;
    when ("Edwards") Nei="EdwardPlusBrkSide";
    when ("BrkSide") Nei="EdwardPlusBrkSide";
    otherwise Nei="Others";
  end;

  if GrLivArea < 4500;
  If GrLivArea > 0;
  if SalePrice > 0;
    logSales=log(SalePrice);
    logLiv=log(GrLivArea);
  if Nei="NAmes" then 
    d2=1;
  else
    d2=0;

  if Nei="Others" then
    d3=1;
  else
    d3=0;

  int1 = d1 * GrLivArea; int2 = d2 * GrLivArea; int3 = d3 * GrLivArea;
  l_int1 = d1 * logLiv; l_int2 = d2 * logLiv; l_int3 = d3 * logLiv;
  centLint1 = (logLiv - 7.2660236) * (d1 - 0.0672154);
  centLint2=(logLiv - 7.2660236) * (d2 - 0.1543210);
  centLint3=(logLiv - 7.2660236) * (d3 - 0.7386831);
run;

proc sort data=trainRed;
  by descending GrLivArea;
run;

proc means data=trainRed;
  var GrLivArea logLiv d2 d3;
run;

proc sgscatter data=trainRed;
  where Nei="Edwards";
  plot SalePrice*GrLivArea / reg;
run;

/******************************************************************************/
/******************************************************************************/
/******************************************************************************/
/********************     Plot Transformed Data       *************************/

proc sgscatter data=trainRed;
  plot logSales*GrLivArea / group=Nei reg;
run;

/* Have Curvature */
proc sgscatter data=trainRed;
  plot SalePrice*logLiv / group=Nei reg;
run;

proc sgscatter data=trainRed;
  plot logSales*logLiv / group=Nei reg;
run;

/******************************************************************************/
/******************************************************************************/
/***********************     Model Building      ******************************/

/* Residual Plot showing possible unequal spread) */
/* Model for Untransformed Data */
proc glm data=trainRed plots=all;
  class Nei(ref="BrkSide");
  model SalePrice=GrLivArea | Nei / solution;
  TITLE "Non Transformed"; 
run;

/************************************/
/* Looks Ok as compared to Previous */
proc glm data=trainRed plots=all;
  class Nei(ref="BrkSide");
  model logSales=GrLivArea | Nei / solution;
  TITLE "Log-Linear Transformed Data"; 
run;

proc reg data=trainRed plots=all;
  model logSales=GrLivArea d1 d2 d3 int1 int2 int3/VIF;
run;

  /************************************/
proc glm data=trainRed plots=all;
  class Nei(ref="EdwardP");
  model logSales=logLiv | Nei / solution;
run;

proc reg data=trainRed plots=all;
  model logSales=logLiv d2 d3 centLint2 centLint3/VIF;
  run;

/******************************************************************************/
/******************************************************************************/
/******************************************************************************/
/******************************************************************************/

/* Create a Heatmap implementation of a correlation matrix */
  ods path work.mystore(update) sashelp.tmplmst(read);

proc template;
  define statgraph corrHeatmap;
    dynamic _Title;
    begingraph;
    entrytitle _Title;
    rangeattrmap name='map';

    /* select a series of colors that represent a "diverging"  */
    /* range of values: stronger on the ends, weaker in middle */
    /* Get ideas from http://colorbrewer.org                   */
    range -1 - 1 / rangecolormodel=(cxD8B365 cxF5F5F5 cx5AB4AC);
    endrangeattrmap;
    rangeattrvar var=r attrvar=r attrmap='map';
    layout overlay / xaxisopts=(display=(line ticks tickvalues)) 
      yaxisopts=(display=(line ticks tickvalues));
    heatmapparm x=x y=y colorresponse=r / xbinaxis=false ybinaxis=false 
      colormodel=THREECOLORRAMP name="heatmap" display=all;
    continuouslegend "heatmap" / orient=vertical location=outside 
      title="Pearson Correlation";
    endlayout;
    endgraph;
  end;
run;

/* Prepare the correlations coeff matrix: Pearson's r method */
%macro prepCorrData(in=, out=);
  /* Run corr matrix for input data, all numeric vars */
  proc corr data=&in. noprint pearson outp=work._tmpCorr vardef=df;
    with SalePrice;
  run;

  /* prep data for heatmap */
  data &out.;
    keep x y r;
    set work._tmpCorr(where=(_TYPE_="CORR"));
    array v{*} _numeric_;
    x=_NAME_;

    do i=dim(v) to 1 by -1;
      y=vname(v(i));
      r=v(i);

      /* creates a diagonally sparse matrix */
      if (i<_n_) then
        r=.;
      output;
    end;
  run;

  proc datasets lib=work nolist nowarn;
    delete _tmpcorr;
  quit;

%mend;

/* Build the graphs */
ods graphics /height=600 width=800 imagemap;
%prepCorrData(in=train0, out=out_r);

proc sgrender data=out_r template=corrHeatmap;
  dynamic _title="Corr matrix for SalePrice";
run;

proc corr data=train0;
  with SalePrice;
run;

data out_r;
  set out_r;
  modr=abs(r);
run;

proc sort data=out_r;
  by descending modr;
run;

proc print data=out_r;
run;

options validvarname=any;

proc corr data=train0 plots=matrix;
  var OverallQual GrLivArea GarageCars GarageArea TotalBsmtSF _1stFlrSF 
    FullBath TotRmsAbvGrd YearBuilt YearRemodAdd GarageYrBlt MasVnrArea 
    Fireplaces;
run;

proc sgscatter data=train0;
  matrix SalePrice OverallQual GrLivArea GarageCars;
run;

proc sgscatter data=train0;
  matrix SalePrice GarageArea TotalBsmtSF _1stFlrSF;
run;

proc sgscatter data=train0;
  matrix SalePrice FullBath TotRmsAbvGrd YearBuilt;
run;

proc sgscatter data=train0;
  matrix SalePrice YearRemodAdd GarageYrBlt MasVnrArea Fireplaces;
run;

/* Test Log Corr */
proc sort data=train0;
  by SalePrice;
run;

proc sgplot data=train0;
  scatter x=Neighborhood y=SalePrice;
run;

proc means data=train0 mean median;
  var saleprice;
  class Neighborhood;
run;

proc glm data=train0;
  class Neighborhood;
  model SalePrice=Neighborhood;
  means Neighborhood/HOVTEST=Levene;
  run;

data train0;
  set train0;
  if Neighborhood = "OldTown" then n1 = 1; else n1 = 0;
  if Neighborhood = "BrkSide" then n2 = 1; else n2 = 0;
  if Neighborhood = "Edwards" then n3 = 1; else n3 = 0;
  if Neighborhood = "SWISU" then n4 = 1; else n4 = 0;
  if Neighborhood = "Sawyer" then n5 = 1; else n5 = 0;
  if Neighborhood = "MeadowV" then n6 = 1; else n6 = 0;
  if Neighborhood = "SawyerW" then n7 = 1; else n7 = 0;
  if Neighborhood = "NWAmes" then n8 = 1; else n8 = 0;
  if Neighborhood = "Mitchel" then n9 = 1; else n9 = 0;
  if Neighborhood = "NAmes" then n10 = 1; else n10 = 0;
  if Neighborhood = "Crawfor" then n11 = 1; else n11 = 0;
  if Neighborhood = "CollgCr" then n12 = 1; else n12 = 0;
  if Neighborhood = "Blueste" then n13 = 1; else n13 = 0;
  if Neighborhood = "NPkVill" then n14 = 1; else n14 = 0;
  if Neighborhood = "ClearCr" then n15 = 1; else n15 = 0;
  if Neighborhood = "Timber" then n16 = 1; else n16 = 0;
  if Neighborhood = "Gilbert" then n17 = 1; else n17 = 0;
  if Neighborhood = "Somerst" then n18 = 1; else n18 = 0;
  if Neighborhood = "NridgHt" then n19 = 1; else n19 = 0;
  if Neighborhood = "Blmngtn" then n20 = 1; else n20 = 0;
  if Neighborhood = "Veenker" then n21 = 1; else n21 = 0;
  if Neighborhood = "StoneBr" then n22 = 1; else n22 = 0;
  if Neighborhood = "NoRidge" then n23 = 1; else n23 = 0;
  if SaleCondition = "Abnorml" then s1 = 1; else s1=0;
  if SaleCondition = "AdjLand" then s2 = 1; else s2=0;
  if SaleCondition = "Alloca" then s3 = 1; else s3=0;
  if SaleCondition = "Family" then s4 = 1; else s4=0;
  if SaleCondition = "Partial" then s5 = 1; else s5=0;
  if GrLivArea < 4000;

  if _n_ = 31 then delete;
  if _n_ = 411 then delete;
  if _n_ = 632 then delete;
  if _n_=1417 then delete;
run;

proc reg data=train0 plots=all;
  model SalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 n18 
    n19 n20 n21 n22 n23/VIF;
run;

data train0;
  set train0;
  LogSalePrice=Log(SalePrice);  /* In */
  LogLotArea=Log(LotArea);  /* In */
  LogOverall=log(OverallQual);  /* out, WO Log is better*/
  LogGrLiving=log(GrLivArea);  /* In, Log is better */
  LogNumCarGar=log(GarageCars);  /* Out, WO Log is better */
  LogGarArea=log(GarageArea);  /* In */
  LogTotalBsmtSF=log(TotalBsmtSF);  /* In */
  Log1stFlrSF=Log(_1stFlrSF);  /* In */
  LogFullBath=log(FullBath);  /* Out, Use WO log */
  LogTotRmsAbvGrd=log(TotRmsAbvGrd);  /* Out, Use WO log */
  /*YearBuilt, YearRemodAdd, GarageYrBlt - No Need to Log */
  LogMasVnrArea=Log(MasVnrArea+1);  /* Potentially In */

  /*LogMasVnrArea = Log(MasVnrArea+1);
  LogBsmtFinSF1 = Log(BsmtFinSF1+1);
  Log1stFlrSF = Log(_1stFlrSF);
  LogPoolArea = Log(PoolArea+1);
  LogMiscVal = Log(MiscVal+1);
  LogGarageArea = Log(GarageArea);
  LogWoodDeckSF = Log(WoodDeckSF+1);
  LogOpenPorchSF = Log(OpenPorchSF+1);
  LogEnclosedPorch = Log(EnclosedPorch+1);
  Log3SsnPorch = Log(_3SsnPorch+1);
  LogScreenPorch = Log(ScreenPorch+1);*/
run;

proc sgscatter data=train0;
  matrix LogSalePrice FullBath LogTotalBsmtSF Log1stFlrSF;
run;

proc reg data=train0 plots=all;
  model LogSalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 
    n18 n19 n20 n21 n22 n23 OverallQual LogLotArea LogGrLiving GarageCars 
    LogGarArea LogTotalBsmtSF Log1stFlrSF FullBath TotRmsAbvGrd YearBuilt 
    YearRemodAdd/vif;
run;

proc reg data=train0 plots=all;
  model LogSalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 
    n18 n19 n20 n21 n22 n23 OverallQual LogLotArea LogGrLiving GarageCars 
    LogGarArea LogTotalBsmtSF Log1stFlrSF FullBath TotRmsAbvGrd YearBuilt 
    YearRemodAdd/vif;
run;

  /* Investigating the Corr among vars, Remove LogFirstSF, LogTotalBsmtSF, TotRmsAbvGrd,LogLotArea area(corr with Gr Living Area),  */
proc sgscatter data=train0;
  matrix OverallQual LogLotArea LogGrLiving GarageCars LogGarArea FullBath 
    YearBuilt YearRemodAdd;
run;

proc sgscatter data=train0;
  plot LogSalePrice *LogGarArea/reg;
run;

proc corr data=train0;
  var LogGarArea GarageCars;
run;

proc reg data=train0 plots(label)=all;
  model LogSalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 
    n18 n19 n20 n21 n22 n23 OverallQual LogGrLiving GarageCars LogGarArea 
    FullBath YearBuilt YearRemodAdd/vif;
run;

  /* Initial Model By Manual */
proc glm data=train0 plots=all;
  class Neighborhood;
  model LogSalePrice=Neighborhood OverallQual LogGrLiving GarageCars LogGarArea 
    FullBath YearBuilt YearRemodAdd;
run;

proc glmselect data=train0;
  class Neighborhood;
  model LogSalePrice=Neighborhood OverallQual LogLotArea LogGrLiving GarageCars 
    LogGarArea LogTotalBsmtSF Log1stFlrSF FullBath TotRmsAbvGrd YearBuilt 
    YearRemodAdd/selection=stepwise stats=ALL;
run;

proc reg data=train0;
  model LogSalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 
    n18 n19 n20 n21 n22 n23 s1 s2 s3 s4 s5 OverallQual LogLotArea LogGrLiving 
    LogGarArea LogTotalBsmtSF Log1stFlrSF TotRmsAbvGrd YearBuilt 
    YearRemodAdd/selection=cp;
  output out=t cookd=cookd;
  plot cp.*np./cmallows=blue vaxis=0 to 15 by 5;
run;

proc reg data=train0 plots(label)=all;
  model LogSalePrice=n1 n2 n3 n4 n5 n6 n7 n8 n9 n10 n11 n12 n13 n14 n15 n16 n17 
    n18 n19 n20 n21 n22 n23 s1 s2 s3 s4 s5 OverallQual LogLotArea LogGrLiving 
    LogGarArea LogTotalBsmtSF YearBuilt YearRemodAdd/vif;
run;

  /* Adding additional Categorial Vars */
proc glmselect data=train0;
  class Neighborhood SaleCondition;
  model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd/selection=stepwise stats=ALL;
run;

proc glmselect data=train0;
  class Neighborhood SaleCondition;
  model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd/selection=stepwise(choose=CV stop=CV) CVDETAILS=ALL;
run;

```
### Test Code

```
/* 
Data Sets: 
- Test0 - Kaggle Test Data Set 
*/

filename CSV URL "https://dl.dropboxusercontent.com/spa/afq05cp80hp4ezn/downloads/public/msds-data/test.csv";

data TEST0;
informat MSZoning $7. ;

/* **************************************** */
/* DATA CLEANUP - use proc means to show missing values*/
/* **************************************** */

title "Test0 - Data Set Missing Values";
proc means data=test0 nmiss n; run;

/* fix missing values */
data test0;
  set test0;
  if missing(LotFrontage) then LotFrontage = 0;
  if missing(MasVnrArea) then MasVnrArea = 0;
  if missing(BsmtFinSF1) then BsmtFinSF1 = 0;
  if missing(BsmtFinSF2) then BsmtFinSF2 = 0;
  if missing(BsmtUnfSF) then BsmtUnfSF = 0;
  if missing(TotalBsmtSF) then TotalBsmtSF = 0;
  if missing(BsmtFullBath) then BsmtFullBath = 0;
  if missing(BsmtHalfBath) then BsmtHalfBath = 0;
  if missing(GarageCars) then GarageCars = 0;
  if missing(GarageArea) then GarageArea = 0;

  /* add log variables from Training model */
  *LogSalePrice=Log(SalePrice);  /* In */
  LogLotArea=Log(LotArea);  /* In */
  *LogOverall=log(OverallQual);  /* out, WO Log is better*/
  LogGrLiving=log(GrLivArea);  /* In, Log is better */
  *LogNumCarGar=log(GarageCars);  /* Out, WO Log is better */
  LogGarArea=log(GarageArea + 1);  /* In */
  LogTotalBsmtSF=log(TotalBsmtSF + 1);  /* In */
  Log1stFlrSF=Log(_1stFlrSF);  /* In */
  *LogFullBath=log(FullBath);  /* Out, Use WO log */
  *LogTotRmsAbvGrd=log(TotRmsAbvGrd);  /* Out, Use WO log */
  /*YearBuilt, YearRemodAdd, GarageYrBlt - No Need to Log */
  LogMasVnrArea=Log(MasVnrArea + 1);  /* Potentially In */

  /*LogMasVnrArea = Log(MasVnrArea+1);
  LogBsmtFinSF1 = Log(BsmtFinSF1+1);
  Log1stFlrSF = Log(_1stFlrSF);
  LogPoolArea = Log(PoolArea+1);
  LogMiscVal = Log(MiscVal+1);
  LogGarageArea = Log(GarageArea);
  LogWoodDeckSF = Log(WoodDeckSF+1);
  LogOpenPorchSF = Log(OpenPorchSF+1);
  LogEnclosedPorch = Log(EnclosedPorch+1);
  Log3SsnPorch = Log(_3SsnPorch+1);
  LogScreenPorch = Log(ScreenPorch+1);*/
run;

title "Test0 - Data Set Fixed Missing Values";
proc means data=test0 nmiss n; run;

title "Test0 - First 5";
proc print data=test0 (obs=5);
run;

```

### Prediction code

```
*FILENAME REFFILE_PREDICT 'c:\temp\kaggle_predict.csv';

/* 
   MSDS6371 - Kaggle Submission

   The Train0 and Test0 data sets have been loaded and cleaned.

   1. load Train0 and Test0 data sets 
      project6371_kaggle_ames_training.sas -> train0
      project6371_kaggle_ames_test.sas     -> test0
   2. run this model, it generates output .csv file. 
   3. login to kaggle portal, navigate to Ames Housing Competition, 
      submit .csv file, You should get score of 0.30871 , lower is better. 
      (take screenshot of Kaggle result if lower for report)
*/

title "Train0 - verify  Missing Values";
proc means data=train0 nmiss n; run;

title "Test0 - verify Missing Values";
proc means data=test0 nmiss n; run;

data test001;
set test0;
SalePrice = .;
LogSalePrice = .;
;

data train001;
set train0 test001;
run;

title "Train0 - First 5";
proc print data=train0 (obs=5);
run;

title "Test0 - First 5";
proc print data=test0 (obs=5);
run;

title "Train001 - First 5";
proc print data=train001 (obs=5);
run;

/* example model
title "Proc GLM on Train001, plots=all. (output predictions, after observation 1460)";
proc glm data=train001 plots=all;
class RoofStyle RoofMatl Exterior1st Exterior2nd MasVnrType;
model SalePrice = RoofStyle RoofMatl Exterior1st Exterior2nd MasVnrType LotArea BedroomAbvGr YearBuilt / cli solution;
output out = results p = Predict;
run;
*/

/* 1. Forward Selection method , stats=PRESS used to get CV Press Score */
/*
proc glmselect data=train001 testdata=test001 plots(stepaxis = number) = (criterionpanel ASEPlot);
model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd
 /selection=Forward(stop=CV) cvmethod=random(5) stats =adjrsq;
output out = results p = Predict;
run;
*/

/* 2. Backward Selection method , stats=PRESS used to get CV Press Score */
/*
proc glmselect data=train001 testdata=test001 plots(stepaxis = number) = (criterionpanel ASEPlot);
model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd
 /selection=Backward(stop=CV) cvmethod=random(5) stats =adjrsq;
output out = results p = Predict;
run;
*/

/* 3. stepwise model */ 
title "Proc GLM on Train001 - 5-Fold Cross Validation. (output predictions, after observation 1460)";
proc glmselect data=train001 testdata=test001 plots(stepaxis = number) = (criterionpanel ASEPlot);
class Neighborhood SaleCondition;
model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd/selection=stepwise(choose=CV stop=CV) CVDETAILS=ALL;
output out = results p = Predict;
run;

/* 4. Forward Selection method , stats=PRESS used to get CV Press Score */
/*
proc glmselect data=train001 testdata=test001 plots(stepaxis = number) = (criterionpanel ASEPlot);
model LogSalePrice=Neighborhood SaleCondition OverallQual LogLotArea 
    LogGrLiving LogGarArea LogTotalBsmtSF YearBuilt 
    YearRemodAdd
 /selection=Forward(select=SL) stats=all;
output out = results p = Predict;
run;
*/

data results0;
set results;
if Predict > 0 then SalePrice = exp(Predict);
if Predict < 0 then SalePrice = 10000;
keep id SalePrice;
where id > 1460;
;

proc means data = results0;
var SalePrice;
run;

PROC EXPORT DATA= results0 
  OUTFILE = 'd:\temp\kaggle_predict_stepwise.csv' replace
  DBMS=CSV REPLACE;
  PUTNAMES=YES;
RUN;

```
## Question 2 Plots

## Image Heat_Map
![Neighborhoods](images/Heat_Map.PNG)

## Image Diagnostics
![Neighborhoods](images/Diagnostics.PNG)

## Image Residual_Plots
![Neighborhoods](images/Residual_Plots.PNG)
